Towards a Machine Learning approach for Aggregate Computing

### 2021-10-05 paper-2021-woa-ac-ml.tex
on introduction

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
\author[1]{Gianluca Aguzzi}[%
orcid=todo,
email=gianluca.aguzzi@unibo.it,
url=https://todo/,
]
\author[2]{Roberto Casadei}[%
orcid=todo,
email=roby.casadei@unibo.it,
url=todo,
]
\author[2]{Mirko Viroli}[%
orcid=todo,
email=mirko.viroli@unibo.it,
url=todo,
]
\address[1]{Alma Mater Studiorum - Università di Bologna,
  Cesena, Italy}
%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  Artificial intelligence is here to stay. 
  In particular, Deep Learning techniques in the last years outperform hand-craft human solutions in different contexts, ranging from computer vision to videogames.
  Currently, though, an all-encompassing Deep learning model does not exist, so researchers explore the possibility to apply these outstanding techniques in different fields.
  One of the most challenging is Collective-Adaptive System (CAS). Here, engineers have to handle large scale, global-to-local behaviour mapping,
  highly environment stochasticity. 
  Historically, researchers favour handling them with self-organisation, by which a collective behaviour emerges from continuously local interaction between simple components.
  A novel technique to specify self-organising behaviour in a composable and functional manner is Aggregate Computing. Even though is promising for expressing high-level collective behaviour, 
  it needs to build low-level blocks that tend to be tricky. 
  On the opposite side, some research lines apply Deep learning (supervised, reinforcement, non-supervised and self-supervised)
  along with CAS. Currently, though, no one gain popularity and solution tend to be application-specific.
  So, we outline a Machine Learning technique that could be applied in Aggregate Computing, making the possibility to combine declarative and automatic approaches that can easily scale in application complexity.
\end{abstract}

@@

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle
\todo[inline, color=red]{max pages : 16}
\section{Introduction}
Aggregate Computing~\cite{DBLP:journals/computer/BealPV15}
\todo[inline]{walkthrogh CAS, why they are hard}
Nowadays, computer engineers have to deal with Collective Adaptive Systems (CASs) due to the increasing interconnected computational devices placed in ever-changing environments.
 Examples of those include robot swarm, smart cities and augmenting crowds.
A CAS is characterised by 
\begin{itemize}
\item a global order that emerges from local node interactions, 
\item  decentralised control, and 
\item adaptive behaviours towards environmental changes to pursuit a collective goal.
\end{itemize}
CAS are notoriously hard to design due as part of complex systems. 
%
So, over the years, in literature, various approaches aim at handling this complexity. 
%
Some of them take inspiration from the natural system composed of social animals -- like bees, fish and ants. 
%
Here, a sort of hive mind arises -- something referred to as swarm intelligence -- born from local interactions between agents. 
%
This kind of behaviour is called self-organisation, namely a global order that emerges from continuous interactions between simple entities.

Initially, researchers try to bring those properties in the computer science world by miming the collective behaviour bottom-up, 
 namely trying to find the single node behaviour to reach a global behaviour, e.g. flocking, foraging, etc.
%
Though, this mapping isn't easily definable -- since we handle complex systems -- therefore novel techniques move the design focus by the ensemble level,
 specifying the system behaviour as a whole.

\todo[inline]{Brief description of declarative approaches, pros and cons}
\todo[inline]{Brief description of machine learning approaches (MARL and GNN for communication), pros and cons}
\todo[inline]{Declare our direction: Combination to declarative and automatic approach}
\todo[inline]{Outline}

@@

\section{Conclusion and Future Work}
\todo[inline]{sharping the model}
\todo[inline]{tune network in order to outperform traditional method}
\todo[inline]{engineering hybrid application (when we need to use ML)}
\todo[inline]{toward online learning?}
%%
%% Define the bibliography file to be used
\bibliography{biblio}

>>>>

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
\author[1]{Gianluca Aguzzi}[%
orcid=todo,
email=gianluca.aguzzi@unibo.it,
url=https://todo/,
]
\author[2]{Roberto Casadei}[%
orcid=todo,
email=roby.casadei@unibo.it,
url=todo,
]
\author[2]{Mirko Viroli}[%
orcid=todo,
email=mirko.viroli@unibo.it,
url=todo,
]
\address[1]{Alma Mater Studiorum - Università di Bologna,
  Cesena, Italy}
%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
Artificial intelligence is here to stay. 
 In particular, Deep Learning techniques in the last years outperform 
 hand-craft human solutions in different contexts, ranging from computer vision to videogames.
%
Currently, though, an all-encompassing Deep learning model does not exist, 
 so researchers explore the possibility to apply these techniques in different fields.

One of the most challenging is Collective-Adaptive System (CAS). % Or we want to use Cyber-Physical Swarm??
 Here, engineers have to handle large scale, global-to-local behaviour mapping,
 highly environment stochasticity. 
%
Historically, researchers favour handling them with self-organisation,
 by which a collective behaviour emerges from continuously local interaction between simple components.
%
A novel technique to specify self-organising behaviour 
 in a composable and functional manner is Aggregate Computing. 
 Even though it is promising for expressing high-level collective behaviour, 
 it needs to build low-level blocks that tend to be tricky.

On the opposite side, some research lines apply Deep learning (supervised, reinforcement, non-supervised and self-supervised)
 along with CAS. Currently, though, no one gain popularity and solution tend to be application-specific.
%
So, we outline a Machine Learning technique that could be applied in Aggregate Computing, 
 making the possibility to combine specification-based and automatic approaches 
 that can easily scale in application complexity and to any system size.
\end{abstract}

@@

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.

%%%%% LEGENDA FOR NOTES %%%%%
%% RED: constraints
%% YELLOW: suggestions
%% CYAN: todos
\maketitle
\todo[inline, color=red]{max pages : 16}
\section{Introduction}
\todo[inline]{walkthrogh CAS, why they are hard}
Nowadays, computer engineers have to deal with Collective Adaptive Systems (CASs)
 due to the increasing interconnected computational devices placed in ever-changing environments.
 Examples of those include robot swarm, smart cities and augmenting crowds.

CASs are characterised by 
\begin{itemize}
\item a global order that emerges from local node interactions, 
\item  decentralised control, and 
\item adaptive behaviours towards environmental changes to pursuit a collective goal.
\end{itemize}
CAS are notoriously hard to design due as part of complex systems. 
%
So, over the years, in literature, various approaches aim at handling this complexity. 
%
Some of them take inspiration from the natural system
 composed of social animals --- like bees, fish and ants. 
%
Here, a sort of hive mind arises -- 
 something referred to as swarm intelligence -- 
born from local interactions between agents. 
%
This kind of behaviour is called self-organisation, 
 namely a global order that emerges from continuous 
 interactions between simple entities.

Initially, researchers try to bring those properties in the computer science world 
 by miming the collective behaviour bottom-up, 
 namely trying to find the right single node behaviour 
 to reach a global behaviour, e.g. flocking, foraging, etc.
%
Though, this mapping isn't easily definable ---
 since we handle complex systems.
%
Ideally we would like to express collective behaviours focusing only at the ensemble level,
 abstracting over non-functional aspect such as the network topology, environment uncertetines and 
 so on.

In this direction, Aggregate Computing is an innovative solution by which 
 developers can express collective self-organising behaviours through a functional program specification.
%
Indeed, an aggregate blueprint consist of a manipulation of \textit{computational fields}, a distributed
 data structure where at each point in the space time is associated the result of a computation
 perform by a node.
%
These manipulation are defined in \textit{Field-calcus} --- the root of Aggregate Computing. 
 This calculus describe a set of minimal constructurs to express any spatiotemporal computation.
%
Thanks to the ``composable'' nature of Aggregate Computing, we find common abstraction 
 in collective behaviour specification and then we incapsulate it in so-called ``building-blocks''.
%
With those, different properties have been proven (such as \textit{eventual consistency}) and various
 application have been build (such as swarm robotics, crowds engineering and IoT like application).
%
Anyway, building block synthesis concerns different stuff related not only to the behaviour itself
 but also about the "ensamble" specifcations, that is some some are orthogonal to the
 beheviour itself. For instance, it is quite difficult to make building block
 that easly scale with the node movements rapidity.
%

So, we claim that Machine Learning can help us in handling this "unseen" situation, leaving only the burden
 to developers of defining the right collective specifications.
%
This lead to an "hibrid" approach where some behaviour part are still express in the of Aggregate Computing
 API, and another part is distilled through learning.

\todo[inline, color=yellow]{I do not know if we can avoid it altogether, but also a brief description of machine learning approaches can be usefull in the paper flow}
\todo[inline, color=cyan]{Outline}

@@

\section{Conclusion and Future Work}
\todo[inline]{sharping the model}
\todo[inline]{tune network in order to outperform traditional method}
\todo[inline]{engineering hybrid application (when we need to use ML)}
\todo[inline]{toward online learning?}
%%
%% Define the bibliography file to be used
\nocite{*}
\bibliography{biblio}
